Type Flattening Obfuscation

### 2020-01-29 paper/tool.tex
add djb

We test outputs of \emph{uCc} against some decompilers, table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of using
$4$ decompilers to decompile binaries generated by \emph{uCc}. The test suite is available at
the repository.

\begin{table}
	\begin{center}
		\caption{Functions and original types}\label{description}
		\begin{tabular}{|l|l|}
			\hline
			Description & Function type \\
			\hline
			identity & \texttt{int id(int)} \\
			\hline
			division & \texttt{int div(short, char)} \\
			\hline
			modular & \texttt{char mod(short, char)} \\
			\hline
			increase $p \leftarrow p + 1$ then dereference & \texttt{int inc\_deref(int *p)} \\
			\hline
			strlen  & \texttt{int slen(char* s)} \\
			\hline
			sum of array $a$ & \texttt{long sum(short *a, int n)} \\
			\hline
			$n$-th fibonaci number (recursive impl.) & \texttt{int fibo(int n)} \\
			\hline
			num. of steps until $n \rightsquigarrow 1$ (Collatz's conj.)  & \texttt{int collatz(int n)} \\
			\hline
			sdbm hash & \texttt{int sdbm(char *s)} \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}
	\begin{adjustwidth}{-6.21in}{-5.95in}
		\begin{center}
			% \centering
			\caption{Type flattening effect}\label{flattening}
			% \begin{adjustbox}{max width=1.5\textwidth}
			\begin{tabular}{|l|l|l|l|l|}
				\hline
				Original type & Hex-Rays & JEB & Ghidra & Snowman \\
				\hline
				\texttt{int id(int)} & \texttt{int64 id(int64)} & \texttt{ulong id(ulong)} &
				\texttt{ulong id(void)} & \texttt{int64 id(int64)} \\
				\hline
				\texttt{short div(short, char)} & \texttt{int64 div(uint64, uint64)} & \texttt{div\_t div(int, int)} &
				\texttt{div\_t div(int, int)} & \texttt{int64\_t div(int64, int64)} \\
				\hline
				\texttt{char mod(short, char)} & \texttt{int64 mod(int64, int64)} & \texttt{ulong mod(ulong, ulong)} &
				\texttt{ulong mod(void)} & \texttt{int64 mod(int64, int64)} \\
				\hline
				\texttt{int inc\_deref(int*)} & \texttt{int64 inc\_deref(int64)} & \texttt{ulong inc\_deref(ulong)} &
				\texttt{ulong inc\_deref(undefined8)} & \texttt{int64 inc\_deref(uint64)} \\
				\hline
				\texttt{int slen(char*)} & \texttt{int64 slen(int64)} & \texttt{ulong slen(ulong)} &
				\texttt{ulong slen(undefined8)} & \texttt{int64 slen(uint64)} \\
				\hline
				\texttt{long sum(short*, int)} & \texttt{int64 sum(int64, uint64)} & \texttt{ulong sum(ulong, ulong)} & \texttt{undefined8 sum(undefined8)} & sn \\
				\hline
				\texttt{int fibo(int)} & \texttt{int64 fibo(uint64)} & \texttt{ulong fibo(ulong)} &
				\texttt{ulong fibo(undefined8)} & \texttt{int64 fibo(int64)} \\
				\hline
				\texttt{int collatz(int)} & \texttt{int64 collatz(int64)} & \texttt{ulong collatz(ulong)} &
				\texttt{ulong collatz(undefined8)} & \texttt{int64 collatz(int64)} \\
				\hline
				\texttt{int sdbm(char*)} & \texttt{int64 sdbm(uint64)} & jeb & \texttt{ulong sdbm(undefined8)} & sn \\
				\hline
			\end{tabular}
		\end{center}
	\end{adjustwidth}
\end{table}

It may worth noting that \emph{uCc} makes no effort in making disassembling or function
boundary detection hard. Indeed, functions in tested binaries are perfectly detectable,
disassemblable and decompilable.

>>>>

We test outputs of \emph{uCc} against some decompilers (the test suite is available at
the repository). Table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of using
$4$ decompilers to decompile binaries generated by \emph{uCc}.
No one between tested decompilers can recognize the precise signature of tested functions.
Basically, any argument is assigned by either \texttt{int64} or \texttt{uint64},
the highest types in the \emph{conversion rank} of integer types, we call this effect \emph{type flattening}.

We may note that if the tested functions are
compiled by, \emph{clang} for example, then their signatures can be recognized precisely
by the decompilers. Last but not least, \emph{uCc} makes no effort in making disassembling or function
boundary detection hard. Indeed, functions in tested binaries are perfectly detectable,
disassemblable and decompilable.

\begin{table}
	\begin{center}
		\caption{Functions and original types}\label{description}
		\begin{tabular}{|l|l|}
			\hline
			Description & Function type \\
			\hline
			identity & \texttt{int id(int)} \\
			\hline
			division & \texttt{int div(short, char)} \\
			\hline
			modular & \texttt{char mod(short, char)} \\
			\hline
			increase $p \leftarrow p + 1$ then dereference & \texttt{int inc\_deref(int *p)} \\
			\hline
			strlen  & \texttt{int slen(char* s)} \\
			\hline
			sdbm hash & \texttt{int sdbm(char *s)} \\
			\hline
			djb2 hash & \texttt{int djb2(char *s)} \\
			\hline
			sum of array $a$ & \texttt{long sum(short *a, int n)} \\
			\hline
			$n$-th fibonaci number (recursive impl.) & \texttt{int fibo(int n)} \\
			\hline
			num. of steps until $n \rightsquigarrow 1$ (Collatz's conj.)  & \texttt{int collatz(int n)} \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}
	\begin{adjustwidth}{-6.21in}{-5.95in}
		\begin{center}
			% \centering
			\caption{Type flattening effect}\label{flattening}
			% \begin{adjustbox}{max width=1.5\textwidth}
			\begin{tabular}{|l|l|l|l|l|}
				\hline
				Original type & Hex-Rays & JEB & Ghidra & Snowman \\
				\hline
				\texttt{int id(int)} & \texttt{int64 id(int64)} & \texttt{ulong id(ulong)} &
				\texttt{ulong id(void)} & \texttt{int64 id(int64)} \\
				\hline
				\texttt{short div(short, char)} & \texttt{int64 div(uint64, uint64)} & \texttt{div\_t div(int, int)} &
				\texttt{div\_t div(int, int)} & \texttt{int64\_t div(int64, int64)} \\
				\hline
				\texttt{char mod(short, char)} & \texttt{int64 mod(int64, int64)} & \texttt{ulong mod(ulong, ulong)} &
				\texttt{ulong mod(void)} & \texttt{int64 mod(int64, int64)} \\
				\hline
				\texttt{int inc\_deref(int*)} & \texttt{int64 inc\_deref(int64)} & \texttt{ulong inc\_deref(ulong)} &
				\texttt{ulong inc\_deref(undefined8)} & \texttt{int64 inc\_deref(uint64)} \\
				\hline
				\texttt{int slen(char*)} & \texttt{int64 slen(int64)} & \texttt{ulong slen(ulong)} &
				\texttt{ulong slen(undefined8)} & \texttt{int64 slen(uint64)} \\
				\hline
				\texttt{int sdbm(char*)} & \texttt{int64 sdbm(uint64)} & \texttt{ulong sdbm(ulong)} &
				\texttt{ulong sdbm(undefined8)} & \texttt{int64 sdbm(uint64)} \\
				\hline
				\texttt{int djb2(char*)} & \texttt{int64 djb2(uint64)} & \texttt{ulong djb2(ulong)} &
				\texttt{ulong djb2(undefined8)} & \texttt{int64 djb2(uint64)} \\
				\hline
				\texttt{long sum(short*, int)} & \texttt{int64 sum(int64, uint64)} &
				\texttt{ulong sum(ulong, ulong)} & \texttt{undefined8 sum(undefined8)} & \texttt{int64 sum(uint64, int64)} \\
				\hline
				\texttt{int fibo(int)} & \texttt{int64 fibo(uint64)} & \texttt{ulong fibo(ulong)} &
				\texttt{ulong fibo(undefined8)} & \texttt{int64 fibo(int64)} \\
				\hline
				\texttt{int collatz(int)} & \texttt{int64 collatz(int64)} & \texttt{ulong collatz(ulong)} &
				\texttt{ulong collatz(undefined8)} & \texttt{int64 collatz(int64)} \\
				\hline
			\end{tabular}
		\end{center}
	\end{adjustwidth}
\end{table}
